---
title: 'Experiments with Sympy'
author: 'fnaufel'
email: 'https://fnaufel.github.io/'
date: '   (v. `r format(Sys.Date(), "%d/%m/%Y")`)'
lang: 'en'

output: 
  # To install these output formats, run
  #   install.packages("devtools")
  #   devtools::install_github("fnaufel/fnaufelRmd")
  fnaufelRmd::html_report:
    []
---

```{r setup, include=FALSE}
# The next command configures MANY things and loads quite a few packages.
# 
# If you want to see what's being done, execute 
# 
#   cat(
#     system.file(
#       "rmarkdown/resources/R/_common_report.R", 
#       package = "fnaufelRmd"
#     )
#   )
# 
# to find out the location of the file. Then open the file.
# 
# If you want to change the configuration, copy the file, edit it, and
# source it instead of the package file. 
# 
# Or simply write your commands here in this code chunk.

source(
  system.file(
    "rmarkdown/resources/R/_common_report.R",
    package = "fnaufelRmd"
  )
)

library(reticulate)
py_run_string('import sympy as sp')
py_run_string('sp.init_printing()')
```


# Printing stuff

## Create stuff

```{python}
print(sp.__version__)
```

```{python}
x, y = sp.symbols('x, y', real=True)
xc, yc = sp.symbols('x_c, y_c', real=True)
rc = sp.symbols('r_c', real=True, positive=True)
```

```{python}
face = sp.Eq((x - xc)**2 + (y - yc)**2, rc**2)
```


## Using variable only

* Ascii math in Python:

```{python}
face
```

```{python}
sp.Integral(x**2, x)
```

* Object in R:

```{r}
py$face
```


## Using print

```{python}
print(face)
```

```{r}
print(py$face)
```


## asis

* Can use LaTeX in python chunk with asis option:

```{python, results='asis'}
print('$$', sp.latex(face), '$$')
```


## Using preview image

```{python}
# Opens viewer window and blocks knitr
# sp.printing.preview(face)
```


## Using asis_output (no interpolation yet)

From https://community.rstudio.com/t/using-sympy-in-rmarkdown/99514

```{r}
mm <- function(x, inline = FALSE) {
  
  py_thing <- rlang::enexpr(x)
  r_name <- rlang::as_label(py_thing)
  py_name <- ifelse(
    startsWith(r_name, 'py$'),
    substring(r_name, 4),
    r_name
  )

  py_run_string(
    paste0(
      'mm_latex = sp.latex(',
      py_name,
      ')'
    )
  )
  
  sep <- ifelse(inline, '$', '$$')
  
  knitr::asis_output(
    paste0(sep, py$mm_latex, sep)
  )

}
```

Display:

```{r}
mm(py$face)
```

Inline: the face has `r mm(py$face, 1)` as its equation.


## Interpolation using glue

See https://glue.tidyverse.org/articles/transformers.html.

```{r}
py_transformer <- function(text, envir) {
  
  py_run_string(
    paste0(
      'mm_latex = sp.latex(',
      text,
      ')'
    )
  )

  py$mm_latex
  
}
```

```{r}
template <- 
r"($$
  \begin{align}
  x_C &= {{xc}} \\ 
  y_C &= {{yc}} \\ 
  r_C &= {{rc}} \\ 
  \text{face} &= {{face}}
  \end{align}
$$)"  

cat(template)
```

```{r}
knitr::asis_output(
  glue::glue(
    template,
    .open = '{{',
    .close = '}}',
    .transformer = py_transformer
  )
)
```


## Production

1. Define function `m('...')`

1. Only argument will be the template string

1. Test in various cases

   a. Decimal point (consult options)
   
   a. Several times in the same chunk
   
   a. Expressions as opposed to variables
   
   a. Inline
   
   a. Inside loop
   
   a. Indentation

```{r}
py_transformer <- function(text, envir) {
  
  decimal <- ifelse(
    getOption('OutDec') == ',',
    '"comma"',
    '"period"'
  )
  
  py_run_string(
    paste0(
      'mm_latex = sp.latex(',
      text,
      ', decimal_separator = ',
      decimal,
      ')'
    )
  )

  py$mm_latex
  
}

m <- function(template, asis = getOption('masis', FALSE)) {
  
  f <- ifelse(asis, cat, knitr::asis_output)

  f(
    glue::glue(
      template,
      .open = '{{',
      .close = '}}',
      .transformer = py_transformer
    )
  )
    
}
```

### Decimal OK

```{python}
sp.var('pi')
```

```{python}
eq = pi / 2
eq = eq.subs(pi, 3.141592)
sp.latex(eq)
```

```{r}
m(
  '$$
  {{eq}}
  $$'
)

options(
  OutDec = '.'
)

m(
  '$$
  {{eq}}
  $$'
)

options(
  OutDec = ','
)
```

Here I can also see that several calls in one chunk just work.


### Expressions OK

* The template string can contain LaTeX expressions:

```{r}
m('$${{xc}}^2 + \\sqrt{ {{rc}} }$$')
```

  Compare 
  
  $$
  x_c^2 + \sqrt{r_c}
  $$

* Python expressions:

```{r}
m('$${{ 2 * pi }}$$')
m('$${{ pi + eq }}$$')
m('$${{ pi + pi }}$$')
```


### Inline R OK

This is an example of inline math: `r m('${{eq + pi}}$')` --- a simple sum.

Can I write $`r m('{{eq + pi}}')`$? Yep.

What if I generate inline [math]{.hl} in a chunk?

```{r}
m('${{eq + pi}}$')
```

I get a paragraph. If it's only math, it's left-aligned.

```{r}
m('I can write text like ${{eq + pi}}$ and it *works*, even the **markdown**.')
```


### Display math

I must have `$$` inside the string that is passed to `m()`.

Wrapping a chunk in `$$` doesn't work:

$$
```{r}
m('{{eq + pi}}')
```
$$


### In a loop

```{r}
for (i in 1:3)
  m(paste0('$${{eq}} + ', i, '$$\n\n'))
```

It doesn't work.

For this I must pass `asis = TRUE` and the chunk must have `results='asis'`:

```{r results='asis'}
for (i in 1:3)
  m(paste0('$${{eq}} + ', i, '$$\n\n'), asis = TRUE)
```

I can set option `masis` to TRUE to make this behavior the default:

```{r results='asis'}
options(masis = TRUE)

for (i in 1:3)
  m(paste0('$${{eq}} + ', i, '$$\n\n'))
```


But now all chunks must have `results='asis'`.

```{r}
options(masis = FALSE)
```


### Can I save the output of `m()`? Yes.

```{r}
output <- m('$${{eq + pi}}$$')
output
```

As `knitr::asis_output` adds a certain class to the string, it still shows asis.


### Indentation OK

1. A list:

    ```{r}
    m('$${{eq + pi}}$$')
    ```

1. Second item.


<div style='height: 300px'></div>
